---
title: "Homework 1"
author: Yiming Zhao
output: github_document
---

This is my solution for HW1

```{r}
library(tidyverse)
data("penguins", package = "palmerpenguins")
```

## Problem 1

###### Create a data frame with the specified elements.

```{r create_df}
prob1_df = 
  tibble(
    samp = rnorm(10),
    samp_gt_0 = samp > 0,
    char_vec = c("a", "b", "c", "d", "e", "f", "g", "h", "i","j"),
    factor_vec = factor(c("low", "low", "low", "mod", "mod", "mod", "mod", "high", "high", "high"))
  )
```

###### Take the mean of each variable in my data frame.

```{r mean_pull_var}
mean(pull(prob1_df, samp))
mean(pull(prob1_df, samp_gt_0))
mean(pull(prob1_df, char_vec))
mean(pull(prob1_df, factor_vec))
```

###### I can take the mean of numbers and logical vectors but not character or factor vectors.


```{r try_numeric, results = 'hide'}
as.numeric(pull(prob1_df, samp))
as.numeric(pull(prob1_df, samp_gt_0))
as.numeric(pull(prob1_df, char_vec))
as.numeric(pull(prob1_df, factor_vec))
```

###### I can convert logical and factor into number, but not character. This is because the logical vectors:True and false, are considered as number 1 and 0 in the R. And Factors are used to categorize the data and store it on multiple levels, the levels can be converted into numbers.
###### These conversions help explain why the mean of numbers and logical vectors can be taken but not character or factor vectors.


```{r convert_logi}
as.numeric(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.factor(pull(prob1_df, samp_gt_0)) * pull(prob1_df, samp)
as.numeric(as.factor(pull(prob1_df, samp_gt_0))) * pull(prob1_df, samp)
```


## Problem 2
###### Write a short description of the penguins dataset:
The variables in penguins dataset includes: `r ls(penguins)`  
The species for penguins in the penguins dataset includes: `r levels(penguins$species)`  
The range year of dataset is `r range(penguins$year, na.rm = FALSE)`  
The range of bill length of penguins in the dataset is `r range(na.omit(penguins$bill_length_mm), na.rm = FALSE)` mm  
The penguins data frame has `r nrow(penguins)` rows.
The penguins data frame has `r ncol(penguins)` columns.  
The mean flipper length for penguins in this data frame is `r mean(na.omit(penguins$flipper_length_mm))` mm

###### Create a scatterplot for bill length vs flipper length, and save the plot as "scatterplot_bill_flipper.png"
```{r penguins_scatterplot}
plot_df = tibble(
  bill_length = penguins$bill_length_mm, 
  flipper_length = penguins$flipper_length_mm,
  species = penguins$species
)

ggplot(plot_df, aes(x = bill_length, y = flipper_length, color = species)) + 
  geom_point() + 
  xlab("bill length (mm)") +
  ylab("flipper length (mm)")

ggsave("scatterplot_bill_flipper.png")
```

